---
title: 'Data Exploration - Game of Thrones : data wrangling and first viz'
author: alexis
date: '2020-04-27'
slug: game-of-thrones-data-wrangling-and-some-viz
categories: []
tags:
  - 'ice and fire'
subtitle: ''
summary: ''
authors: []
lastmod: '2020-04-27T11:39:20+02:00'
featured: yes
image:
  caption: ''
  focal_point: ''
  preview_only: yes
projects: []
---

I recently stumbled upon a great data source on Game of Thrones, it lead me to spend (lose) some (a lot of) time playing around with it.
The data comes from here : https://github.com/jeffreylancaster/game-of-thrones with some very neat visualisations already.

I was very excited, first thing to do get the data into tidy format :
I import the data from github, at the moment I am intersted in scene details : characters involved, location, duration etc...

```{r setup, include=FALSE}
library(tidyverse)
library(here)

```


```{r eval=FALSE}
library(tidyverse)
library(jsonlite)
library(here)
library(lubridate)
library(here)

got <- fromJSON("https://raw.githubusercontent.com/jeffreylancaster/game-of-thrones/master/data/episodes.json")

scenes <- got$episodes %>%
  as_tibble() %>% 
  select(-episodeLink,
         -episodeAirDate,
         -episodeDescription,
         -openingSequenceLocations) %>% 
  unnest(scenes) %>% 
  group_by(seasonNum, episodeNum, episodeTitle) %>% 
  mutate(sceneNum = 1,
         sceneNum = cumsum(sceneNum)) %>% 
  unnest(characters) %>% 
  select(seasonNum,
         episodeNum,
         episodeTitle,
         sceneNum,
         sceneStart,
         sceneEnd,
         location,
         subLocation,
         name) %>% 
  ungroup() %>% 
  mutate(sceneStart = hms(sceneStart),
         sceneEnd = hms(sceneEnd),
         time = seconds(sceneEnd - sceneStart) %>% as.numeric())


write_csv(scenes, here("static/inputs/got/data_prep.csv"))

```


```{r message=FALSE}
#Let's load some more packages for visualisation
library(patchwork)
#for got themed color palette
library(gameofthrones)
library(hrbrthemes)


extrafont::loadfonts(device = "win")

got <- read_csv(here("static/inputs/got/data_prep.csv"))
```

I used hrbrthemes because the default theme is great, I particularly like the font size default which is really convenient. 
I don't think a legend is needed as colors are in order.

```{r, fig.height=12}

got %>% 
  group_by(name, seasonNum) %>%
  summarise(time = sum(time) / 60) %>% 
  mutate(total = sum(time)) %>% 
  filter(total > 60) %>% 
  ggplot(aes(y = fct_reorder(name, total), 
             x = time, 
             fill = fct_rev(as_factor(seasonNum)))) +
  geom_col() +
  geom_vline(xintercept = 0) +
  scale_fill_got_d(direction = -1) +
  labs(y = NULL,
       x = "Time on screen in minutes",
       title = "Game of Thrones main characters:\nTime on screen per season",
       subtitle = "Characters that appear at least one hour on screen, colored by season") +
  scale_x_continuous(expand = expansion(mult = c(0.01,0.05))) +
  theme_ipsum(grid = "X") +
  theme(legend.position = "none",
        plot.title = element_text(family = "Cinzel"),
        
        plot.title.position = "plot")


ggsave(here("static/featured.png"), height = 10, width = 7, dpi = 600, type = "cairo")

```

