---
title: 'Creating a choropleth map with Voronoï polygons'
author: admin
date: '2020-05-05'
slug: creating-a-choropleth-map-with-voronoï-polygons-a-peek-into-the-importance-of-projection
categories:
  - maps
tags: []
subtitle: ''
summary: 'A peek into the importance of choosing a projected coordinate system'
authors: []
lastmod: '2020-05-05T15:51:22+02:00'
featured: yes
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
output:
  blogdown::html_page:
    dev: "svg"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.retina = 2,
                      message = FALSE,
                      warning = FALSE)
```

I learned R by giving myself small challenges to overcome. When I started playing around with spatial data, one particular challenge helped me to understand projections better.
What I wanted to do,  is to make a map that showed everywhere in France what is the neareast football club (in Ligue 1) and to color the map based on the total area in which each club is the neareast in France.

I will first show you how to do this "properly". Which mean using **st_transform** on the data to convert the coordinates system to a Lambert Projection (Lambert93 which is the official projection for France).

Then I will show you the first result I got, when I did not projected the data and instead used **st_voronoi** directly on the longitude/lattitude data.

Finally, a little bonus : a raster map of distance to nearest club everywhere in France.

## Setting up
### Loading packages

```{r}
#Usual suspects
library(tidyverse)
library(glue)
library(here)
#To work with geodata
library(sf)
#For geodata of the cities
library(httr)
library(jsonlite)
#For plotting
library(ggrepel)

#Extra fonts for maps
extrafont::loadfonts(device = "win")
```

### theme_map() 

I have a theme_map() function which comes from another project. Let's use that.

```{r}
#theme map function from other project
theme_map <- function(...) {
  theme_minimal() +
    theme(
      text = element_text(family = "Tahoma", color = "#22211d"),
      #remove all axes
      axis.line = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      axis.title = element_blank(),
      #remove grid
      panel.grid = element_blank(),
      #background colors
      plot.background = element_rect(color = NA,
                                     fill = "#f5f5f2"),
      panel.background = element_rect(color = NA,
                                      fill = "#f5f5f2"),
      legend.background = element_rect(color = NA,
                                       fill = "#f5f5f2"),
      #borders and margins
      plot.margin = unit(c(0.5, 0.5, 0.2, 0.5), "cm"),
      panel.border = element_blank(),
      panel.spacing = unit(c(-0.1, 0.2, 0.2, 0.2), "cm"),
      #titles
      legend.title = element_text(size = 11),
      legend.text = element_text(size = 9, hjust = 0),
      plot.title = element_text(size = 15, hjust = 0.5),
      plot.subtitle = element_text(size = 10, hjust = 0.5,
                                   margin = margin(b = -0.1,
                                                   t = -0.1,
                                                   l = 2,
                                                   unit = "cm"),
                                   debug = FALSE),
      #captions
      plot.caption = element_text(size = 7,
                                  hjust = 1,
                                  margin = margin(t = 0.2,
                                                  b = 0,
                                                  unit = "cm"),
                                  color = "#939184"),
      ...
    )
}
```

### Getting the geodata

We will be using the national address database, which is a convenient way to get localisation in France.
I also use a France simplified geojson : that you can get here : [boundary.now](https://haoliangyu.github.io/boundary.now/).

```{r, message=FALSE}
#clubs
clubs <- tibble(city =
                  c(
                    "Lille", 
                    "Lens", 
                    "Paris", 
                    "Reims",
                    "Metz",
                    "Strasbourg",
                    "Dijon",
                    "Rennes",
                    "Angers",
                    "Lorient",
                    "Brest",
                    "Nantes",
                    "Bordeaux",
                    "Saint-Étienne",
                    "Lyon",
                    "Nice",
                    "Marseille",
                    "Montpellier",
                    "Nîmes"
                  ))



get_location <- function(city) {
  city %>% 
    paste0("https://api-adresse.data.gouv.fr/search/?q=",.) %>%
    GET() %>%
    pluck("content") %>%
    rawToChar() %>%
    fromJSON() %>%
    pluck("features") %>% 
    flatten() %>%
    slice(1) %>% 
    pluck("geometry.coordinates", 1) %>% 
    set_names(c("LON", "LAT")) %>% 
    list()
}


clubs_geo <- clubs %>% 
  rowwise() %>% 
  mutate(coords = get_location({city})) %>% 
  unnest_wider(coords) %>% 
  # I manually add Monaco which is not in France
  bind_rows(tibble(city = "Monaco", LON = 7.414793, LAT = 43.728353)) %>% 
  st_as_sf(coords = c("LON", "LAT")) %>% 
  st_set_crs(4326) %>% 
  st_transform(2154)
  
clubs_geo <- clubs_geo %>% 
  bind_cols(as_tibble(
    st_coordinates(clubs_geo)
  ))



#France
france <- st_read(
  here("static/inputs/ligue 1 voronoi/metropole-version-simplifiee.geojson")
  ) %>%
  st_transform(2154) %>%
  st_cast("POLYGON") %>%
  mutate(area = st_area(geometry)) %>%
  filter(as.numeric(area) == max(as.numeric(area)))



```

## Creating Voronoi polygons

```{r}
#make voronoi ploygons
voronoi <- clubs_geo %>%
  #to make voronoi polygons you need every point in the same feature
  st_union() %>%
  st_voronoi() %>%
  st_cast() %>%
  st_sf() %>% 
  st_join(clubs_geo, join = st_contains) %>% 
  st_intersection(france) %>% 
  mutate(area = st_area(geometry))
```


## Plotting

```{r, fig.height=7.45}
#plot description
description <- glue(
  "For each club area in which it is the
  closest club in France"
  )


m1 <- ggplot(voronoi) +
  geom_sf(aes(fill = as.numeric(area) / 1000000)) +
  geom_sf(data = clubs_geo) +
  geom_label_repel(data = clubs_geo,
                   aes(label = city,
                       x = X,
                       y = Y)) +
  scale_fill_distiller(direction = 1,name = "Surface",
                       labels = scales::label_comma(big.mark = " ",
                                                    suffix = " km²"),
                       breaks = c(10000, 30000, 50000, 70000), 
                       guide = guide_colorbar(barwidth = 0.5,
                                              frame.colour = "black")) +
  annotate("text", x=-Inf, y=Inf, 
           label = description,
           hjust = 0, vjust = 1) +
  theme_map() +
  theme(legend.position = c(0.1,0.2),
        plot.title = element_text(hjust = 0),
        plot.subtitle = element_text(hjust = 0),
        plot.background = element_rect(color = "black")) +
  labs(title = "Territorial domination of Ligue 1 clubs",
       caption = "@_Alexis_69_")

# ggsave(plot = m1, here("static/plots/ligue1_voronoi/map.png"),
#        dpi = "retina",
#        width = 7, height = 7.4)

m1


```

## What if we did not change the projection ?

The code is exactly the same, except we don't use **st_transform**, therefore did not include it. But you can see the warning thrown : "st_voronoi does not correctly triangulate longitude/latitude data".  
And if you look carefully at the map you can tell it is not exactly right.

```{r echo=FALSE, message=FALSE, fig.height=7.15, warning=TRUE}

clubs_geo_bad <- clubs %>% 
  rowwise() %>% 
  mutate(coords = get_location({city})) %>% 
  unnest_wider(coords) %>% 
  bind_rows(tibble(city = "Monaco", LON = 7.414793, LAT = 43.728353)) %>% 
  st_as_sf(coords = c("LON", "LAT")) %>% 
  st_set_crs(4326)

clubs_geo_bad <- clubs_geo_bad %>% 
  bind_cols(as_tibble(
    st_coordinates(clubs_geo_bad)
  ))

#France
france_bad <- st_read(
  here("static/inputs/ligue 1 voronoi/metropole-version-simplifiee.geojson")
  ) %>%
  st_cast("POLYGON") %>%
  mutate(area = st_area(geometry)) %>%
  filter(as.numeric(area) == max(as.numeric(area)))


voronoi_bad <- clubs_geo_bad %>%
  st_union() %>%
  st_voronoi() %>%
  st_cast() %>%
  st_sf() %>% 
  st_join(clubs_geo_bad, join = st_contains) %>% 
  st_intersection(france_bad) %>% 
  mutate(area = st_area(geometry))

ggplot(voronoi_bad) +
  geom_sf(aes(fill = as.numeric(area) / 1000000)) +
  geom_sf(data = clubs_geo_bad) +
  geom_label_repel(data = clubs_geo_bad,
                   aes(label = city,
                       x = X,
                       y = Y)) +
  scale_fill_distiller(direction = 1,name = "Surface",
                       labels = scales::label_comma(big.mark = " ",
                                                    suffix = " km²"),
                       breaks = c(10000, 30000, 50000, 70000), 
                       guide = guide_colorbar(barwidth = 0.5,
                                              frame.colour = "black")) +
  annotate("text", x=-Inf, y=Inf, 
           label = glue("At first, the map may seem fine, but if you pay attention 
                        the line are not where you would want them
                        to be..."),
           hjust = 0, vjust = 1) +
  theme_map() +
  theme(legend.position = c(0.1,0.2),
        plot.title = element_text(hjust = 0),
        plot.subtitle = element_text(hjust = 0),
        plot.background = element_rect(color = "black")) +
  labs(title = "Bad Map ! ",
       caption = "@_Alexis_69_")

```



## Bonus : Raster map

```{r, fig.height=7.4}
library(raster)


r <- raster(france, ncol = 500, nrow = 500)
distances <- distanceFromPoints(r, clubs_geo) %>% 
  mask(france) %>% 
  as.data.frame(xy = TRUE) %>% 
  na.omit() %>% 
  as_tibble() %>% 
  mutate(distance = layer / 1000) %>% 
  dplyr::select(-layer) %>% 
  mutate(bins = case_when(
           distance < 50 ~ "<50 km",
           distance <= 100 ~ "50 à 100 km",
           distance <= 150 ~ "100 à 150 km",
           distance <= 200 ~ "150 à 200 km",
           distance >200 ~ ">200 km"),
         bins = fct_reorder(bins, distance)) 
 
description_2 <- "Distance in kms to the nearest Ligue 1 football club"

ggplot() +
  geom_raster(data = distances, aes(fill = bins,
                                    x = x,
                                    y = y),
              interpolate = TRUE) +
  scale_fill_brewer(name = NULL) +
  geom_sf(data = clubs_geo) +
  geom_sf(data = france,
          fill = "transparent") +
  geom_label_repel(data = clubs_geo,
                   aes(x = X,
                       y = Y,
                       label = city)) +
  annotate("text", x = -Inf, y = Inf, 
           label = description_2, 
           hjust = 0, vjust = 1) +
  theme_map() +
  theme(legend.position = c(0.1,0.2),
        plot.title = element_text(hjust = 0),
        legend.key = element_rect(color = "black"),
        plot.background = element_rect(color = "black")) +
  labs(title = "La diagonale du vide du football français",
       caption = "@_Alexis_69_")

```

