---
title: 'Creating a choropleth map with Voronoï polygons'
author: admin
date: '2020-05-05'
slug: creating-a-choropleth-map-with-voronoï-polygons-a-peek-into-the-importance-of-projection
categories:
  - maps
tags: []
subtitle: ''
summary: 'A peek into the importance of choosing the right projection'
authors: []
lastmod: '2020-05-05T15:51:22+02:00'
featured: yes
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
output:
  blogdown::html_page:
    dev: "svg"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.retina = 2,
                      message = FALSE,
                      warning = FALSE)
```

In this article I want two things : how to build a choropleth map with Vornoï polygon using sf.
The map I will be drawing is the area in which each Ligue 1 football team is the nearest in France.
First, I will use the recommanded legal projection for France, which is Lambert projection (crs : 2154). Then I will show the results using the WGS64 projection which is used to share geospatial data.

## Loading packages

```{r}
#Usual suspects
library(tidyverse)
library(glue)
library(here)
#To work with geodata
library(sf)
#For geodata of the cities
library(httr)
library(jsonlite)
#For plotting
library(ggrepel)

#Extra fonts for maps
extrafont::loadfonts(device = "win")
```

## theme_map()

I have a theme_map() function which comes from another project. Let's use that !

```{r}
#theme map function from other project
theme_map <- function(...) {
  theme_minimal() +
    theme(
      text = element_text(family = "Tahoma", color = "#22211d"),
      #remove all axes
      axis.line = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      axis.title = element_blank(),
      #remove grid
      panel.grid = element_blank(),
      #background colors
      plot.background = element_rect(color = NA,
                                     fill = "#f5f5f2"),
      panel.background = element_rect(color = NA,
                                      fill = "#f5f5f2"),
      legend.background = element_rect(color = NA,
                                       fill = "#f5f5f2"),
      #borders and margins
      plot.margin = unit(c(0.5, 0.5, 0.2, 0.5), "cm"),
      panel.border = element_blank(),
      panel.spacing = unit(c(-0.1, 0.2, 0.2, 0.2), "cm"),
      #titles
      legend.title = element_text(size = 11),
      legend.text = element_text(size = 9, hjust = 0),
      plot.title = element_text(size = 15, hjust = 0.5),
      plot.subtitle = element_text(size = 10, hjust = 0.5,
                                   margin = margin(b = -0.1,
                                                   t = -0.1,
                                                   l = 2,
                                                   unit = "cm"),
                                   debug = FALSE),
      #captions
      plot.caption = element_text(size = 7,
                                  hjust = 1,
                                  margin = margin(t = 0.2,
                                                  b = 0,
                                                  unit = "cm"),
                                  color = "#939184"),
      ...
    )
}
```

## Getting the geodata

We will be using the national address database, which is a convenient way to get localisation in France.
I also use a France simplified geojson : that you can get here : [boundaty.now](https://haoliangyu.github.io/boundary.now/).

```{r}
#clubs
clubs <- tibble(city =
                  c(
                    "Lille", 
                    "Lens", 
                    "Paris", 
                    "Reims",
                    "Metz",
                    "Strasbourg",
                    "Dijon",
                    "Rennes",
                    "Angers",
                    "Lorient",
                    "Brest",
                    "Nantes",
                    "Bordeaux",
                    "Saint-Étienne",
                    "Lyon",
                    "Nice",
                    "Marseille",
                    "Montpellier",
                    "Nîmes"
                  ))



get_location <- function(city) {
  city %>% 
    paste0("https://api-adresse.data.gouv.fr/search/?q=",.) %>%
    GET() %>%
    pluck("content") %>%
    rawToChar() %>%
    fromJSON() %>%
    pluck("features") %>% 
    flatten() %>%
    slice(1) %>% 
    pluck("geometry.coordinates", 1) %>% 
    set_names(c("LON", "LAT")) %>% 
    list()
}


clubs_geo <- clubs %>% 
  rowwise() %>% 
  mutate(coords = get_location({city})) %>% 
  unnest_wider(coords) %>% 
  # I manually add Monaco which is not in France
  bind_rows(tibble(city = "Monaco", LON = 7.414793, LAT = 43.728353)) %>% 
  st_as_sf(coords = c("LON", "LAT")) %>% 
  st_set_crs(4326) %>% 
  st_transform(2154)
  
clubs_geo <- clubs_geo %>% 
  bind_cols(as_tibble(
    st_coordinates(clubs_geo)
  ))



#France
france <- st_read(
  here("static/inputs/ligue 1 voronoi/metropole-version-simplifiee.geojson")
  ) %>%
  st_transform(2154) %>%
  st_cast("POLYGON") %>%
  mutate(area = st_area(geometry)) %>%
  filter(as.numeric(area) == max(as.numeric(area)))



```


```{r}
#make voronoi ploygons
voronoi <- clubs_geo %>%
  st_union() %>%
  st_voronoi() %>%
  st_cast() %>%
  st_sf() %>% 
  st_join(clubs_geo, join = st_contains) %>% 
  st_intersection(france) %>% 
  mutate(area = st_area(geometry))
```


```{r, fig.height=7.45}
#plot description
description <- glue(
  "For each club area in which it is the
  closest club in France"
  )


m1 <- ggplot(voronoi) +
  geom_sf(aes(fill = as.numeric(area) / 1000000)) +
  geom_sf(data = clubs_geo) +
  geom_label_repel(data = clubs_geo,
                   aes(label = city,
                       x = X,
                       y = Y)) +
  scale_fill_distiller(direction = 1,name = "Surface",
                       labels = scales::label_comma(big.mark = " ",
                                                    suffix = " km²"),
                       breaks = c(10000, 30000, 50000, 70000), 
                       guide = guide_colorbar(barwidth = 0.5,
                                              frame.colour = "black")) +
  annotate("text", x=-Inf, y=Inf, 
           label = description,
           hjust = 0, vjust = 1) +
  theme_map() +
  theme(legend.position = c(0.1,0.2),
        plot.title = element_text(hjust = 0),
        plot.subtitle = element_text(hjust = 0),
        plot.background = element_rect(color = "black")) +
  labs(title = "Territorial domination of Ligue 1 clubs",
       caption = "@_Alexis_69_")

# ggsave(plot = m1, here("static/plots/ligue1_voronoi/map.png"),
#        dpi = "retina",
#        width = 7, height = 7.4)

m1


```

## What if we did not change the projection

The code is exactly the same, except we don't use **st_transform**, therefore I not include the code but you can see the warning thrown : "st_voronoi does not correctly triangulate longitude/latitude data".  
And if you look carefully at the map you can tell it is not exactly right.

```{r echo=FALSE, message=FALSE, fig.height=7.15, warning=TRUE}

clubs_geo_bad <- clubs %>% 
  rowwise() %>% 
  mutate(coords = get_location({city})) %>% 
  unnest_wider(coords) %>% 
  bind_rows(tibble(city = "Monaco", LON = 7.414793, LAT = 43.728353)) %>% 
  st_as_sf(coords = c("LON", "LAT")) %>% 
  st_set_crs(4326)

clubs_geo_bad <- clubs_geo_bad %>% 
  bind_cols(as_tibble(
    st_coordinates(clubs_geo_bad)
  ))

#France
france_bad <- st_read(
  here("static/inputs/ligue 1 voronoi/metropole-version-simplifiee.geojson")
  ) %>%
  st_cast("POLYGON") %>%
  mutate(area = st_area(geometry)) %>%
  filter(as.numeric(area) == max(as.numeric(area)))


voronoi_bad <- clubs_geo_bad %>%
  st_union() %>%
  st_voronoi() %>%
  st_cast() %>%
  st_sf() %>% 
  st_join(clubs_geo_bad, join = st_contains) %>% 
  st_intersection(france_bad) %>% 
  mutate(area = st_area(geometry))

ggplot(voronoi_bad) +
  geom_sf(aes(fill = as.numeric(area) / 1000000)) +
  geom_sf(data = clubs_geo_bad) +
  geom_label_repel(data = clubs_geo_bad,
                   aes(label = city,
                       x = X,
                       y = Y)) +
  scale_fill_distiller(direction = 1,name = "Surface",
                       labels = scales::label_comma(big.mark = " ",
                                                    suffix = " km²"),
                       breaks = c(10000, 30000, 50000, 70000), 
                       guide = guide_colorbar(barwidth = 0.5,
                                              frame.colour = "black")) +
  annotate("text", x=-Inf, y=Inf, 
           label = glue("At first, the map may seem fine, but if you pay attention 
                        the line are not where you would want them
                        to be..."),
           hjust = 0, vjust = 1) +
  theme_map() +
  theme(legend.position = c(0.1,0.2),
        plot.title = element_text(hjust = 0),
        plot.subtitle = element_text(hjust = 0),
        plot.background = element_rect(color = "black")) +
  labs(title = "Bad Map ! ",
       caption = "@_Alexis_69_")

```



## Bonus : Raster map

```{r, fig.height=7.4}
library(raster)


r <- raster(france, ncol = 500, nrow = 500)
distances <- distanceFromPoints(r, clubs_geo) %>% 
  mask(france) %>% 
  as.data.frame(xy = TRUE) %>% 
  na.omit() %>% 
  as_tibble() %>% 
  mutate(distance = layer / 1000) %>% 
  dplyr::select(-layer) %>% 
  mutate(bins = case_when(
           distance < 50 ~ "<50 km",
           distance <= 100 ~ "50 à 100 km",
           distance <= 150 ~ "100 à 150 km",
           distance <= 200 ~ "150 à 200 km",
           distance >200 ~ ">200 km"),
         bins = fct_reorder(bins, distance)) 
 
description_2 <- "Distance in kms to the nearest Ligue 1 football club"

ggplot() +
  geom_raster(data = distances, aes(fill = bins,
                                    x = x,
                                    y = y),
              interpolate = TRUE) +
  scale_fill_brewer(name = NULL) +
  geom_sf(data = clubs_geo) +
  geom_sf(data = france,
          fill = "transparent") +
  geom_label_repel(data = clubs_geo,
                   aes(x = X,
                       y = Y,
                       label = city)) +
  annotate("text", x = -Inf, y = Inf, 
           label = description_2, 
           hjust = 0, vjust = 1) +
  theme_map() +
  theme(legend.position = c(0.1,0.2),
        plot.title = element_text(hjust = 0),
        legend.key = element_rect(color = "black"),
        plot.background = element_rect(color = "black")) +
  labs(title = "La diagonale du vide du football français",
       caption = "@_Alexis_69_")

```

